maze-dataset

### 2025-04-08 docs/paper/paper.tex
examples, docs, notebooks link

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
    similar to the outputs of Németh
    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to that used in
    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
  Wu, 2023}), but provides over 5.8 million unique formats for
  converting mazes to a text stream, detailed in
  \autoref{tokenized-output-formats}.
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
  this reliably.
\item
  Storing mazes as images is not only difficult to work with, but also
  inefficient. We use a highly efficient method detailed in
  \autoref{implementation}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
  2008}) appear to simply no longer be accessible.
\end{itemize}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
\NormalTok{)}
\CommentTok{\# create a config}
\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
    \CommentTok{\# (optional) algorithm{-}specific parameters}
\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
    \CommentTok{\# (optional) many options for restricting start/end points}
\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
\NormalTok{)}
\CommentTok{\# create a dataset}
\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
\NormalTok{  cfg, }\CommentTok{\# pass the config}
\NormalTok{  ..., }\CommentTok{\# other options such as whether to load from disk,}
       \CommentTok{\# generate in parallel, etc.}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

>>>>

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
    similar to the outputs of Németh
    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to
    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
  Wu, 2023}), but provides over 5.8 million unique formats for
  converting mazes to a text stream, detailed in
  \autoref{tokenized-output-formats}.
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
  this reliably.
\item
  Storing mazes as images is not only difficult to work with, but also
  inefficient. We use a highly efficient method detailed in
  \autoref{implementation}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
  2008}) appear to simply no longer be accessible.
\end{itemize}

@@

We direct readers to our
\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
and
\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
for more information.

@@

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{from}\NormalTok{ maze\_dataset }\ImportTok{import}\NormalTok{ (}
\NormalTok{  MazeDataset, MazeDatasetConfig, LatticeMazeGenerators}
\NormalTok{)}
\CommentTok{\# create a config}
\NormalTok{cfg: MazeDatasetConfig }\OperatorTok{=}\NormalTok{ MazeDatasetConfig(}
\NormalTok{    name}\OperatorTok{=}\StringTok{"example"}\NormalTok{, }\CommentTok{\# names need not be unique}
\NormalTok{    grid\_n}\OperatorTok{=}\DecValTok{3}\NormalTok{,   }\CommentTok{\# size of the maze}
\NormalTok{    n\_mazes}\OperatorTok{=}\DecValTok{32}\NormalTok{, }\CommentTok{\# number of mazes in the dataset}
\NormalTok{    maze\_ctor}\OperatorTok{=}\NormalTok{LatticeMazeGenerators.gen\_dfs, }\CommentTok{\# many algorithms available}
    \CommentTok{\# (optional) algorithm{-}specific parameters}
\NormalTok{    maze\_ctor\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"do\_forks"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\}, }
    \CommentTok{\# (optional) many options for restricting start/end points}
\NormalTok{    endpoint\_kwargs}\OperatorTok{=}\NormalTok{\{}\StringTok{"deadend\_start"}\NormalTok{: }\VariableTok{True}\NormalTok{, ...\},}
\NormalTok{)}
\CommentTok{\# create a dataset}
\NormalTok{dataset: MazeDataset }\OperatorTok{=}\NormalTok{ MazeDataset.from\_config(}
\NormalTok{  cfg, }\CommentTok{\# pass the config}
\NormalTok{  ..., }\CommentTok{\# other options for disk loading, parallelization, etc.}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}
